---
title: "D607 Project 2 - Tidy Data"
author: "Kyle Gilde"
date: "Mar. 4, 2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#Chapter 4 Foundations for Inference
Practice: 4.3, 4.13, 4.23, 4.25, 4.39, 4.47
Graded: 4.4, 4.14, 4.24, 4.26, 4.34, 4.40, 4.48

#####Load Packages
```{r setup} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      tidy=F,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))
#https://cran.r-project.org/web/packages/prettydoc/vignettes/
#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

load.packages <- c("prettydoc","knitr")

ipak <- function(pkg){
    #FUNCTION SOURCE: https://gist.github.com/stevenworthington/3178163
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}

ipak(load.packages)
```

####4.4 Heights of adults. 
Researchers studying anthropometry collected body girth measurements and skeletal diameter measurements, as well as age, weight, height and gender, for 507 physically active individuals. The histogram below shows the sample distribution of heights in centimeters.

(a) What is the point estimate for the average height of active individuals? **171.1**

What about the median? **170.3**

(b) What is the point estimate for the standard deviation of the heights of active individuals? **9.4**

What about the IQR?
```{r}
177.8 - 163.8
```

(c) Is a person who is 1m 80cm (180 cm) tall considered unusually tall? **This person is not unusually tall because his/her height has a z-score of less than one and is within one standard deviation of the mean. For a normal distribution, approximately 34% of the sample is at least this extreme.**
And is a person who is 1m 55cm (155cm) considered unusually short? Explain your reasoning. **This person is more unusally short because her z-score of -1.71 is further away from the mean. Only 9% of the sample is at least this extreme**

```{r}
v1 <- 180
v2 <- 155
xbar <- 171.1
s <- 9.4

#person 1 zscore
(v1 - xbar)/s
pnorm(v1, xbar, s, lower.tail = F) * 2

#person 2 zscore
(v2 - xbar)/s
pnorm(v2, xbar, s) * 2
```


(d) The researchers take another random sample of physically active individuals. Would you expect the mean and the standard deviation of this new sample to be the ones given above? Explain your reasoning.
**No, they would not exactly match. Both point estimates could be found on the sampling distribution.**
(e) The sample means obtained are point estimates for the mean height of all active individuals, if the sample of individuals is equivalent to a simple random sample. What measure do we use to quantify the variability of such an estimate)? Compute this quantity using the data from the original sample under the condition that the data are a simple random sample. **The standard error**

```{r }
s / sqrt(507)
```

####4.14 Thanksgiving spending
Part I. The 2009 holiday retail season, which kicked off on November 27, 2009 (the day after Thanksgiving), had been marked by somewhat lower self-reported consumer spending than was seen during the comparable period in 2008. To get an estimate of consumer spending, 436 randomly sampled American adults were surveyed. Daily consumer spending for the six-day period after Thanksgiving, spanning the Black Friday weekend and Cyber Monday, averaged \$84.71. A 95% confidence interval based on this sample is (\$80.31, \$89.11).

Determine whether the following statements are true or false, and explain your reasoning.

(a) We are 95% confident that the average spending of these 436 American adults is between $80.31 and \$89.11. **False. Confidence intervals pertain to the population parameter. We are 95% confident that the average post-Thanksgiving spending for American adults is between \$80.31 and \$89.11.**

(b) This confidence interval is not valid since the distribution of spending in the sample is right skewed.
**False. Because we have a large sample size, we can be more lenient with the not-strongly-skewed condition.**

(c) 95% of random samples have a sample mean between \$80.31 and \$89.11. **False. If repeated samples were taken and the 95% confidence interval was computed for each sample, 95% of the intervals would contain the population mean.**


(d) We are 95% confident that the average spending of all American adults is between \$80.31 and \$89.11. **True. This is the correct interpretation of the CI**

(e) A 90% confidence interval would be narrower than the 95% confidence interval since we don't need to be as sure about our estimate. **True. The critical value would be less, which would create a smaller margin of error when it's multipled by the SE.**

(f) In order to decrease the margin of error of a 95% confidence interval to a third of what it is now, we would need to use a sample 3 times larger. **False. Because the SE and therefore the ME is divided by the square root of the sample size, we would need to take a sample 9 times larger.**
```{r}
n <- 436
xbar <- 84.71 
ci1 <- 80.31 
ci2 <- 89.11
me <- xbar - ci1
zstar <- 1.96
se <- me / zstar
s <- se * sqrt(n)

n2 <- n * 9
me2 <- zstar * s/sqrt(n2)
me/me2
```
(g) The margin of error is 4.4. **True. The ME is calculated by adding and substracting it to the point estimate.**

```{r }
xbar - ci1
```

####4.24 Gifted children, Part I. 
Researchers investigating characteristics of gifted children collected data from schools in a large city on a random sample of thirty-six children who were identified as gifted children soon after they reached the age of four. The following histogram shows the distribution of the ages (in months) at which these children first counted to 10 successfully. Also provided are some sample statistics.



(a) Are conditions for inference satisfied?
**Yes, the observations are independent since they are from a random sample of less than 10% of the population. The sample size is greater than or equal to 30, and the histogram doesn't a strong skew**

(b) Suppose you read online that children first count to 10 successfully when they are 32 months old, on average. Perform a hypothesis test to evaluate if these data provide convincing evidence that the average age at which gifted children first count to 10 successfully is less than the general average of 32 months. Use a significance level of 0.10.
$H_{0}: \mu_{gifted}=32$ 
$H_{A}:  \mu_{gifted}<32$


```{r }
n <- 36
min_months <- 21
xbar <- 30.69
s <- 4.31
max_months <- 39

sign_level <- .10

se <- s/sqrt(n)
cv <- qnorm(.05)


```

(c) Interpret the p-value in context of the hypothesis test and the data.

(d) Calculate a 90% confidence interval for the average age at which gifted children first count to 10 successfully.

(e) Do your results from the hypothesis test and the confidence interval agree? Explain.

##Which airline & city had the best on-time arrival rate?
```{r}


```

##Which airline had the better on-time arrival rate?
```{r analysis, tidy=F}

```

##Let's rank the cities by their overall on-time arrival rates.
```{r}

```

