---
title: "D606 Homework Chapter 3"
author: "Kyle Gilde"
date: "Feb. 19, 2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

##OpenIntro Statistics

Practice: 3.1 (see normalPlot), 3.3, 3.17 (use qqnormsim from lab 3), 3.21, 3.37, 3.41
Graded: 3.2 (see normalPlot), 3.4, 3.18 (use qqnormsim from lab 3), 3.22, 3.38, 3.42

##Load Packages
```{r setup, results='hide'} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      tidy=TRUE,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))

#https://cran.r-project.org/web/packages/prettydoc/vignettes/
#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

load.packages <- c("DATA606")


ipak <- function(pkg){
    #FUNCTION SOURCE: https://gist.github.com/stevenworthington/3178163
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}
ipak(load.packages)

```



##3.2 Area under the curve, Part II. 

What percent of a standard normal distribution N(?? = 0,= 1) is found in each region? Be sure to draw a graph.

(a) Z > -1.13 (b) Z < 0.18 (c) Z > 8 (d) |Z| < 0.5
```{r 3.2}
normalPlot(bounds=c(-1.13, Inf))
normalPlot(bounds=c(-Inf, .18))
normalPlot(bounds=c(8, Inf))
normalPlot(bounds=c(-.5,.5))
```
##3.4 Triathlon times, Part I. 

In triathlons, it is common for racers to be placed into age and gender groups. Friends Leo and Mary both completed the Hermosa Beach Triathlon, where Leo competed in the Men, Ages 30 - 34 group while Mary competed in the Women, Ages 25 - 29 group. Leo completed the race in 1:22:28 (4948 seconds), while Mary completed the race in 1:31:53 (5513 seconds). Obviously Leo finished faster, but they are curious about how they did within their respective groups. Can you help them? Here is some information on the performance 
of their groups:

. The finishing times of the Men, Ages 30 - 34 group has a mean of 4313 seconds with a standard deviation of 583 seconds.

. The finishing times of the Women, Ages 25 - 29 group has a mean of 5261 seconds with a standard deviation of 807 seconds.

. The distributions of finishing times for both groups are approximately Normal.

Remember: a better performance corresponds to a faster finish.

(a) Write down the short-hand for these two normal distributions. 

(b) What are the Z-scores for Leo's and Mary's finishing times? What do these Z-scores tell you?

```{r}
leo <- 4948
mary <- 5513

male.mu <- 4313
male.sd <- 583

female.mu <- 5261
female.sd <- 807

leo.z <- (leo - male.mu)/male.sd
mary.z <- (mary - female.mu)/female.sd

leo.z; mary.z

print("Mary's z-score is relatively closer to her group's mean than Leo's.")
```


(c) Did Leo or Mary rank better in their respective groups? Explain your reasoning.
**Since her z-score is less than Leo's **


(d) What percent of the triathletes did Leo finish faster than in his group?

(e) What percent of the triathletes did Mary finish faster than in her group?

(f) If the distributions of finishing times are not nearly normal, would your answers to parts

(b) - (e) change? Explain your reasoning.

```{r }



```


