---
title: "D606 Chapter 5 & 6"
author: "Kyle Gilde"
date: "Mar. 14, 2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---


#Load Packages
```{r setup} 
knitr::opts_chunk$set(warning=FALSE, 
                      message=FALSE,
                      tidy=F,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))
#https://cran.r-project.org/web/packages/prettydoc/vignettes/
#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

#create vector with all needed packages
load_packages <- c("prettydoc", "stringr", "knitr")

#see if we need to install any of them
install_load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

install_load(load_packages)
#CODE SOURCE DOCUMENTATION: https://gist.github.com/stevenworthington/3178163
```

#Chapter 5 - Inference for Numerical Data
Practice: 5.5, 5.13, 5.19, 5.31, 5.45
Graded: 5.6, 5.14, 5.20, 5.32, 5.48



####5.6 Working backwards, Part II. 
A 90% confidence interval for a population mean is (65, 77). The population distribution is approximately normal and the population standard deviation is unknown. This confidence interval is based on a simple random sample of 25 observations. Calculate the sample mean, the margin of error, and the sample standard deviation.

```{r}
#givens
ci <- .9
ci_values <- c(65, 77)
n <- 25
df <- n - 1

#calcs
me <- (ci_values[2] - ci_values[1]) / 2
xbar <- ci_values[1] + me
cv <- abs(qt((1 - ci) / 2, df))
          
          
se <- me / cv
s <- se * sqrt(n)

#answers
xbar
me
s
```


####5.13 Car insurance savings. 
A market researcher wants to evaluate car insurance savings at a competing company. Based on past studies he is assuming that the standard deviation of savings is \$100. He wants to collect data such that he can get a margin of error of no more than $10 at a 95% confidence level. How large of a sample should he collect?

```{r}
#givens
sigma <- 100
ci <- .95
me <- 10

#calcs
cv <- abs(qnorm((1 - ci) / 2))
se <- me / cv
n <- (sigma / se) ^ 2

ceiling(n)
```

####5.14 SAT scores. 
SAT scores of students at an Ivy League college are distributed with a standard deviation of 250 points. Two statistics students, Raina and Luke, want to estimate the average SAT score of students at this college as part of a class project. They want their margin of error to be no more than 25 points.

(a) Raina wants to use a 90% confidence interval. How large a sample should she collect?
```{r}
#givens
sigma <- 250
ci <- .9
me <- 25

#calcs
cv <- abs(qnorm((1 - ci) / 2))
se <- me / cv
n <- (sigma / se) ^ 2

ceiling(n)
```

(b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size,
determine whether his sample should be larger or smaller than Raina's, and explain your
reasoning.

**Given a maximum ME of 25, the sample size is going to have to be larger. With a CI of .99, the critical value (CV) is going to increase. Since ME is SE * CV , the SE is going to need to be smaller to still get an ME of 25. Since SE is the SD divided by the square root of n, in order to get a smaller SE, the sample size will need larger.**


(c) Calculate the minimum required sample size for Luke.
```{r}
#givens
sigma <- 250
ci <- .99
me <- 25

#calcs
cv <- abs(qnorm((1 - ci) / 2))
se <- me / cv
n <- (sigma / se) ^ 2

ceiling(n)
```

####5.19 Global warming, Part I. 
Is there strong evidence of global warming? Let's consider a small scale example, comparing how temperatures have changed in the US from 1968 to 2008. The daily high temperature reading on January 1 was collected in 1968 and 2008 for 51 randomly selected locations in the continental US. Then the difference between the two readings (temperature in 2008 - temperature in 1968) was calculated for each of the 51 different locations. The average of these 51 values was 1.1 degrees with a standard deviation of 4.9 degrees. We are interested in determining whether these data provide strong evidence of temperature warming in the continental 
US.
(a) Is there a relationship between the observations collected in 1968 and 2008? Or are the observations in the two groups independent? Explain.
**No, they have a special correspondence since the same locations were used in both samples. Consequently, we should use the t-distribution.**
(b) Write hypotheses for this research in symbols and in words.

$H_{0}: \mu_{diff}=0$ 
$H_{A}:  \mu_{gifted}>0$

(c) Check the conditions required to complete this test.
(d) Calculate the test statistic and find the p-value.
(e) What do you conclude? Interpret your conclusion in context.
(f) What type of error might we have made? Explain in context what the error means.
(g) Based on the results of this hypothesis test, would you expect a confidence interval for the
average difference between the temperature measurements from 1968 and 2008 to include 0?
Explain your reasoning.

####5.20 High School and Beyond, Part I. 
The National Center of Education Statistics conducted a survey of high school seniors, collecting test data on reading, writing, and several other subjects. Here we examine a simple random sample of 200 students from this survey. Side-by-side box plots of reading and writing scores as well as a histogram of the differences in scores are shown below.

(a) Is there a clear difference in the average reading and writing scores?
(b) Are the reading and writing scores of each student independent of each other?
(c) Create hypotheses appropriate for the following research question: is there an evident difference
in the average scores of students in the reading and writing exam?
(d) Check the conditions required to complete this test.
(e) The average observed difference in scores is ¯xreadwrite
= 0.545,
and the standard deviation
of the differences is 8.887 points. Do these data provide convincing evidence of a difference
between the average scores on the two exams?
(f) What type of error might we have made? Explain what the error means in the context of the
application.
(g) Based on the results of this hypothesis test, would you expect a confidence interval for the
average difference between the reading and writing scores to include 0? Explain your reasoning.

5.31 Chicken diet and weight, Part I. Chicken farming is a multi-billion dollar industry,
and any methods that increase the growth rate of young chicks can reduce consumer costs while
increasing company profits, possibly by millions of dollars. An experiment was conducted to
measure and compare the effectiveness of various feed supplements on the growth rate of chickens.
Newly hatched chicks were randomly allocated into six groups, and each group was given a different
feed supplement. Below are some summary statistics from this data set along with box plots
showing the distribution of weights by feed type.41 Weight (in grams)
casein horsebean linseed meatmeal soybean sunflower

Mean SD n
casein 323.58 64.43 12
horsebean 160.20 38.63 10
linseed 218.75 52.24 12
meatmeal 276.91 64.90 11
soybean 246.43 54.13 14
sunflower 328.92 48.84 12
(a) Describe the distributions of weights of chickens that were fed linseed and horsebean.
(b) Do these data provide strong evidence that the average weights of chickens that were fed
linseed and horsebean are different? Use a 5% significance level.
(c) What type of error might we have committed? Explain.
(d) Would your conclusion change if we used ff = 0.01?


5.32 Fuel eciency of manual and automatic cars, Part I. 

Each year the US Environmental
Protection Agency (EPA) releases fuel economy data on cars manufactured in that year.
Below are summary statistics on fuel eciency
(in miles/gallon) from random samples of cars with
manual and automatic transmissions manufactured in 2012. Do these data provide strong evidence
of a difference between the average fuel eciency
of cars with manual and automatic transmissions
in terms of their average city mileage? Assume that conditions for inference are satisfied.42

5.45 Coffee, depression, and physical activity. 
Caffeine is the world's most widely used stimulant, with approximately 80% consumed in the form of coffee. Participants in a study investigating the relationship between coffee consumption and exercise were asked to report the number of hours they spent per week on moderate (e.g., brisk walking) and vigorous (e.g., strenuous sports and jogging) exercise. Based on these data the researchers estimated the total hours of metabolic equivalent tasks (MET) per week, a value always greater than 0. The table below gives summary statistics of MET for women in this study based on the amount of coffee consumed.46

Caffeinated coffee consumption

(a) Write the hypotheses for evaluating if the average physical activity level varies among the
different levels of coffee consumption.
(b) Check conditions and describe any assumptions you must make to proceed with the test.
(c) Below is part of the output associated with this test. Fill in the empty cells.
Df Sum Sq Mean Sq F value Pr(>F)
(d) What is the conclusion of the test?

5.48 Work hours and education. 
The General Social Survey collects data on demographics, education, and work, among many other characteristics of US residents.47 Using ANOVA, we can consider educational attainment levels for all 1,172 respondents at once. Below are the distributions of hours worked by educational attainment and relevant summary statistics that will be helpful in carrying out this analysis.
Educational attainment 

Less than HS HS Jr Coll Bachelor's Graduate Total
Mean 38.67 39.6 41.39 42.55 40.85 40.45
SD 15.81 14.97 18.1 13.62 15.51 15.17
n 121 546 97 253 155 1,172
Hours worked per week
Less than HS HS Jr Coll Bachelor's Graduate
0
20
40
60
80
(a) Write hypotheses for evaluating whether the average number of hours worked varies across the
five groups.
(b) Check conditions and describe any assumptions you must make to proceed with the test.
(c) Below is part of the output associated with this test. Fill in the empty cells.
Df Sum Sq Mean Sq F value Pr(>F)
degree XXXXX XXXXX 501.54 XXXXX 0.0682
Residuals XXXXX 267,382 XXXXX
Total XXXXX XXXXX

#Chapter 6 - Inference for Categorical Data
Practice: 6.5, 6.11, 6.27, 6.43, 6.47
Graded: 6.6, 6.12, 6.20, 6.28, 6.44, 6.48

6.5 Prop 19 in California. In a 2010 Survey USA poll, 70% of the 119 respondents between
the ages of 18 and 34 said they would vote in the 2010 general election for Prop 19, which would
change California law to legalize marijuana and allow it to be regulated and taxed. At a 95%
confidence level, this sample has an 8% margin of error. Based on this information, determine if
the following statements are true or false, and explain your reasoning.38
(a) We are 95% confident that between 62% and 78% of the California voters in this sample
support Prop 19.
(b) We are 95% confident that between 62% and 78% of all California voters between the ages of
18 and 34 support Prop 19.
(c) If we considered many random samples of 119 California voters between the ages of 18 and
34, and we calculated 95% confidence intervals for each, 95% of them will include the true
population proportion of 18-34 year old Californians who support Prop 19.
(d) In order to decrease the margin of error to 4%, we would need to quadruple (multiply by 4)
the sample size.
(e) Based on this confidence interval, there is sucient
evidence to conclude that a majority of
California voters between the ages of 18 and 34 support Prop 19.

6.6 2010 Healthcare Law. On June 28, 2012 the U.S. Supreme Court upheld the much debated
2010 healthcare law, declaring it constitutional. A Gallup poll released the day after this decision
indicates that 46% of 1,012 Americans agree with this decision. At a 95% confidence level, this
sample has a 3% margin of error. Based on this information, determine if the following statements
are true or false, and explain your reasoning.39
(a) We are 95% confident that between 43% and 49% of Americans in this sample support the
decision of the U.S. Supreme Court on the 2010 healthcare law.
(b) We are 95% confident that between 43% and 49% of Americans support the decision of the
U.S. Supreme Court on the 2010 healthcare law.
(c) If we considered many random samples of 1,012 Americans, and we calculated the sample
proportions of those who support the decision of the U.S. Supreme Court, 95% of those sample
proportions will be between 43% and 49%.
(d) The margin of error at a 90% confidence level would be higher than 3%

6.11 Study abroad. A survey on 1,509 high school seniors who took the SAT and who completed
an optional web survey between April 25 and April 30, 2007 shows that 55% of high school seniors
are fairly certain that they will participate in a study abroad program in college.43
(a) Is this sample a representative sample from the population of all high school seniors in the
US? Explain your reasoning.
(b) Let's suppose the conditions for inference are met. Even if your answer to part (a) indicated
that this approach would not be reliable, this analysis may still be interesting to carry out
(though not report). Construct a 90% confidence interval for the proportion of high school
seniors (of those who took the SAT) who are fairly certain they will participate in a study
abroad program in college, and interpret this interval in context.
(c) What does "90% confidence" mean?
(d) Based on this interval, would it be appropriate to claim that the majority of high school seniors
are fairly certain that they will participate in a study abroad program in college?
42Gallup World, More Than One in 10 "Suffering" Worldwide, data collected throughout 2011. 43studentPOLL, College-Bound Students' Interests in Study Abroad and Other International Learning
Activities, January 2008.



6.12 Legalization of marijuana, Part I. The 2010 General Social Survey asked 1,259 US residents:
"Do you think the use of marijuana should be made legal, or not?" 48% of the respondents
said it should be made legal.44
(a) Is 48% a sample statistic or a population parameter? Explain.
(b) Construct a 95% confidence interval for the proportion of US residents who think marijuana
should be made legal, and interpret it in the context of the data.
(c) A critic points out that this 95% confidence interval is only accurate if the statistic follows a
normal distribution, or if the normal model is a good approximation. Is this true for these
data? Explain.
(d) A news piece on this survey's findings states, "Majority of Americans think marijuana should
be legalized." Based on your confidence interval, is this news piece's statement justified?
6.13 Public option, Part I. A Washington Post article from 2009 reported that "support
for a government-run health-care plan to compete with private insurers has rebounded from its
summertime lows and wins clear majority support from the public." More specifically, the article
says "seven in 10 Democrats back the plan, while almost nine in 10 Republicans oppose it. Independents
divide 52 percent against, 42 percent in favor of the legislation." (6% responded with
"other".) There were 819 Democrats, 566 Republicans and 783 Independents surveyed.45
(a) A political pundit on TV claims that a majority of Independents oppose the health care public
option plan. Do these data provide strong evidence to support this statement?
(b) Would you expect a confidence interval for the proportion of Independents who oppose the
public option plan to include 0.5? Explain.

6.19 College smokers. We are interested in estimating the proportion of students at a university
who smoke. Out of a random sample of 200 students from this university, 40 students smoke.
(a) Calculate a 95% confidence interval for the proportion of students at this university who smoke,
and interpret this interval in context. (Reminder: Check conditions.)
(b) If we wanted the margin of error to be no larger than 2% at a 95% confidence level for the
proportion of students who smoke, how big of a sample would we need?

6.20 Legalize Marijuana, Part II. As discussed in Exercise 6.12, the 2010 General Social
Survey reported a sample where about 48% of US residents thought marijuana should be made
legal. If we wanted to limit the margin of error of a 95% confidence interval to 2%, about how
many Americans would we need to survey ?

6.27 Public Option, Part III. Exercise 6.13 presents the results of a poll evaluating support
for the health care public option plan in 2009. 70% of 819 Democrats and 42% of 783 Independents
support the public option.
(a) Calculate a 95% confidence interval for the difference between (pD
pI ) and interpret it in
this context. We have already checked conditions for you.
(b) True or false: If we had picked a random Democrat and a random Independent at the time
of this poll, it is more likely that the Democrat would support the public option than the
Independent.


6.28 Sleep deprivation, CA vs. OR, Part I. According to a report on sleep deprivation by the
Centers for Disease Control and Prevention, the proportion of California residents who reported
insucient
rest or sleep during each of the preceding 30 days is 8.0%, while this proportion is
8.8% for Oregon residents. These data are based on simple random samples of 11,545 California
and 4,691 Oregon residents. Calculate a 95% confidence interval for the difference between the
proportions of Californians and Oregonians who are sleep deprived and interpret it in context of
the data.53

6.43 Rock-paper-scissors. Rock-paper-scissors is a hand game played by two or more people
where players choose to sign either rock, paper, or scissors with their hands. For your statistics
class project, you want to evaluate whether players choose between these three options randomly,
or if certain options are favored above others. You ask two friends to play rock-paper-scissors and
count the times each option is played. The following table summarizes the data:
Rock Paper Scissors
43 21 35
Use these data to evaluate whether players choose between these three options randomly, or if
certain options are favored above others. Make sure to clearly outline each step of your analysis,
and interpret your results in context of the data and the research question.


6.44 Barking deer. Microhabitat factors associated with forage and bed sites of barking deer
in Hainan Island, China were examined from 2001 to 2002. In this region woods make up 4.8%
of the land, cultivated grass plot makes up 14.7%, and deciduous forests makes up 39.6%. Of the
426 sites where the deer forage, 4 were categorized as woods, 16 as cultivated grassplot, and 61 as
deciduous forests. The table below summarizes these data.62
Woods Cultivated grassplot Deciduous forests Other Total
4 16 61 345 426
(a) Write the hypotheses for testing if barking deer prefer to forage
in certain habitats over others.
(b) What type of test can we use to answer this research question?
(c) Check if the assumptions and conditions required for this test
are satisfied.
(d) Do these data provide convincing evidence that barking deer prefer
to forage in certain habitats over others? Conduct an appropriate
hypothesis test to answer this research question.

6.47 Offshore drilling, Part III. The table below summarizes a data set we first encountered
in Exercise 6.29 that examines the responses of a random sample of college graduates and nongraduates
on the topic of oil drilling. Complete a chi-square test for these data to check whether
there is a statistically significant difference in responses from college graduates and non-graduates.
College Grad
Yes No
Support 154 132
Oppose 180 126
Do not know 104 131
Total 438 389


6.48 Coffee and Depression. Researchers conducted a study investigating the relationship
between caffeinated coffee consumption and risk of depression in women. They collected data
on 50,739 women free of depression symptoms at the start of the study in the year 1996, and
these women were followed through 2006. The researchers used questionnaires to collect data on
caffeinated coffee consumption, asked each individual about physician-diagnosed depression, and
also asked about the use of antidepressants. The table below shows the distribution of incidences
of depression by amount of caffeinated coffee consumption.63
Caffeinated coffee consumption
??? 1 2-6 1 2-3
4
cup/week cups/week cup/day cups/day cups/day Total
Clinical Yes 670 373 905 564 95 2,607
depression No 11,545 6,244 16,329 11,726 2,288 48,132
Total 12,215 6,617 17,234 12,290 2,383 50,739
(a) What type of test is appropriate for evaluating if there is an association between coffee intake
and depression?
(b) Write the hypotheses for the test you identified in part (a).
(c) Calculate the overall proportion of women who do and do not suffer from depression.
(d) Identify the expected count for the highlighted cell, and calculate the contribution of this cell
to the test statistic, i.e. (Observed
Expected)
2/Expected.
(e) The test statistic is 2
= 20.93. What is the p-value?
(f) What is the conclusion of the hypothesis test?
(g) One of the authors of this study was quoted on the NYTimes as saying it was "too early to
recommend that women load up on extra coffee" based on just this study.64 Do you agree with
this statement? Explain your reasoning.

#Practice
####5.5 Working backwards, Part I. 
A 95% confidence interval for a population mean, µ, is given as (18.985, 21.015). This confidence interval is based on a simple random sample of 36 observations. Calculate the sample mean and standard deviation. Assume that all conditions necessary for inference are satisfied. Use the t-distribution in any calculations.

```{r}
#givens
ci <- .95
ci_values <- c(18.985, 21.015)
n <- 36

#calcs
me <- (ci_values[2] - ci_values[1]) / 2
xbar <- ci_values[1] + me
cv <- abs(qt((1 - ci) / 2, n - 1))
se <- me / cv
s <- se * sqrt(n)

xbar
s
```

